---
title: "Bullying Network"
author: "Aurora Maria Tumminello, Lucia Hrovatin"
output: html_document
---


# Importing data

## Data pre-processing 
A pre-processing of the original dataset has been necessary. Thus, the original Excel table contained two sheets. Whereas a first attempt involved the function **read_excel(path, sheet = 2)** and implied the network's misreading, a manual split solved the deadlock. Therefore, separate CSV files have been created: 
<li> *bullying_attributes.csv*: containing the attributes of the network </li><br>
<li> *bullying_network.csv* : containing the actual matrix (i.e., network) </li><br> 

```{r}
library(tidyverse)
library(sna)

# Import of network data
net = as.matrix(read.csv("bullying_network.csv", 
                         stringsAsFactors=FALSE, row.names=1))
net = as.network(net, directed=T)
summary(net)
```


```{r}
# Import of attributes
attr = read.csv("bullying_attributes.csv")
```


# Drawing a network

*Draw the network using the Fruchterman-Reingold layout with names (but without attributes).*

```{r}
par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,           # do not allow nodes to be "jittered"
      edge.col="grey70",  # set color of ties
      vertex.col="cornflowerblue",   # set color of nodes
      displaylabels=T,    # indicate that labels should be included
      label.pos=1,        # indicate that labels should be given below points
      label.cex=.7)       # indicate the size of the labels (1 is default)
```

------------------------------------------------------------------------

*Next draw the network again using the Fruchterman-Reingold layout without names, but now with all 3 attributes (gender, grade and ethnicity) included using in some way a combination of color, shape and size of nodes.*

```{r}
# SIZE: Grade
# SHAPE: Gender
# COLOR: Ethnicity

colors = c('#fff0f3', '#ffccd5', '#ffb3c1', '#ff8fa3', '#ff758f', 
           '#ff4d6d', '#c9184a', '#a4133c', '#800f2f', '#590d22',
           "#d8f3dc","#b7e4c7","#95d5b2","#74c69d","#52b788",
           "#40916c", "#2d6a4f","#245741","#1b4332","#081c15"
)

shapes = c(3,4,50)

par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,           # do not allow nodes to be "jittered"
      edge.col="grey70",  # set color of ties
      vertex.col=colors[attr$Gender*10+attr$Grade+1],   # set color of nodes
      vertex.cex = 2,
      vertex.rot = 45,
      vertex.sides= shapes[attr$Ethnicity])

```

------------------------------------------------------------------------

*Create a legend for the attributes (this does not have to be with R, but needs to clarify how the different attributes were represented in the network).*

```{r}
color = c("#2a9d8f", "#e9c46a", "#e76f51")
  
par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,           # do not allow nodes to be "jittered"
      edge.col="grey70",  # set color of ties
      vertex.col=color[attr$Ethnicity],   # set color of nodes
      vertex.cex = log(attr$Grade),
      vertex.sides= (attr$Gender-1)*47+4,
      displaylabels=T,    # indicate that labels should be included
      label.pos=1,        # indicate that labels should be given below points
      label.cex=.7)       # indicate the size of the labels (1 is default)
legend("topright",
       legend = c("Ethnicity"," 1st gen immigrants", " 2st gen immigrants", " Italian",
                  " ", "Gender", " Male", " Female",
                  " ", "Grade", "0", "3", "7", "10"),
       col = c("white",color[1],color[2],color[3],
               "white","white","black","black",
               "white", "white", "black", "black", "black", "black"), 
       bty = "n", 
       pch =c(15,15,15,15,15,15,19,18),
       pt.cex = c(0,2.1,2.1, 2.1,
                  0,0,2.1,2.1,
                  0, 0, 1, log(3), log(7), log(10)), 
       cex = 1, 
       text.col = "grey10", 
       horiz = F , 
       inset = c(0.01))

```

or

```{r}
par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,           # do not allow nodes to be "jittered"
      edge.col="grey70",  # set color of ties
      vertex.col=colors[attr$Gender*10+attr$Grade+1],   # set color of nodes
      vertex.cex = 1.5,
      vertex.rot = 45,
      vertex.sides= shapes[attr$Ethnicity],
      displaylabels=T,    # indicate that labels should be included
      label.pos=1,        # indicate that labels should be given below points
      label.cex=.7)       # indicate the size of the labels (1 is default))
legend("topleft",
       legend = c("Ethnicity"," 1st gen immigrants", " 2st gen immigrants", " Italian",
                  " ", "Gender", " Male", " Female",
                  " ", "Grade"),
       col = c("white","black","black", "black",
               "white","white","#52b788","#ff4d6d",
               "white", "white"), 
       bty = "n", 
       pch =c(15,17,15,19,
              15,15,15,15,
              15, 15),
       pt.cex = c(0,2.1,2.1, 2.2,
                  0,0,2.1,2.1,
                  0,0), 
       cex = 1, 
       text.col = "grey10", 
       horiz = F , 
       inset = c(0.01))

library(itsadug)
pal = colorRampPalette(colors[1:10])
gradientLegend(valRange=c(0,10), 
               color = pal(10),
               pos=c(0.05,0.1,.1,0.5), side=4, length = 0.5,
               inside = T)

pal = colorRampPalette(colors[11:20])
gradientLegend(valRange=c(0,10), 
               color = pal(10),
               pos=c(0.15,0.1,.2,0.5), side=4, length = 0.5,
               inside = T)
```

## Centrality

*Calculate in and out degree centrality for each node and interpret the results:*

-   Who is the most central when it comes to indegree?

-   Provide details of the meaning of the value (taking into account that the network is "who do you bully?").

-   Make sure to discuss the reference points for the indegree measure. How does it compare to the theoretical minimum and maximum value?

```{r}
# From how many people are you bullied? 
indegree = sna::degree(net, gmode="digraph", cmode = "indegree")
names = get.vertex.attribute(net, "vertex.names")
people = data.frame(indegree, names)


par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,
      edge.col="grey70",  # set color of ties
      vertex.col=color[attr$Ethnicity],   # set color of nodes
      vertex.cex = log(indegree),
      vertex.sides= (attr$Gender-1)*47+4,
      displaylabels=T,    # indicate that labels should be included
      label.pos=1,        # indicate that labels should be given below points
      label.cex=.7)       # indicate the size of the labels (1 is default)


```

-   Who is the most central when it comes to outdegree?

-   Provide details of the meaning of the value (taking into account that the network is "who do you bully?").

-   Make sure to discuss the reference points for the outdegree measure. How does it compare to the theoretical minimum and maximum value?

```{r}
# How many people do you bully? 
outdegree = sna::degree(net, gmode="digraph", cmode = "outdegree")
par(mar=c(0,0,0,0)) # Margin deletion
gplot(net,
      gmode="digraph",    # directed network
      jitter=F,
      edge.col="grey70",  # set color of ties
      vertex.col=color[attr$Ethnicity],   # set color of nodes
      vertex.cex = log(outdegree)+1,
      vertex.sides= (attr$Gender-1)*47+4,
      displaylabels=T,    # indicate that labels should be included
      label.pos=1,        # indicate that labels should be given below points
      label.cex=.7)       # indicate the size of the labels (1 is default)
```

------------------------------------------------------------------------

*Let's assume we hypothesize that the grade of students might be impacted by how many others they are being bullied by. Calculate the correlation between the indegree and the grade of the kid.*

-   What is the correlation value?
-   What does it mean substantively?

```{r}

```

------------------------------------------------------------------------

*Perform a significance test using a permutation-based approach for the correlation as discussed in class. What does the significance test do and what do you conclude?*

```{r}

```


# Centralization and reciprocity

*Calculate the indegree and outdegree centralization for this network using the approach discussed in class and interpret. Make sure to show how you came to these answers.*

-   What do you conclude substantively (taking into account that the network is "who do you bully?")?
-   Make sure to discuss the reference points for the measure. What is the minimum and maximum value?

```{r}

```


------------------------------------------------------------------------

*Calculate the arc-based reciprocity index for this network and interpret. Make sure to show how you came to this answer.*

-   What do you conclude substantively (taking into account that the network is "who do you bully?")?
-   Make sure to discuss the reference points for the measure. What is the minimum, expected and maximum value?

```{r}

```

